---
title: "Untitled"
author: "Rick Fontenot"
date: "4/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Load Libraries
```{r load-packages, include=FALSE}
library(dplyr)
library(tidyverse)
library(caret)
library(DataExplorer)
library(gplots)
library(graphics)
library(corrplot)
library(olsrr)
library(ggpubr)
library(rstatix)
library(dplyr)
library(tidyverse)
library(visdat)
library(GGally)
library(usmap)
library(mice)
library(VIM)
library(plotly)
library(caret)
library(e1071)
library(class)
library(maps)
library(mapproj)
library(stringr)
library(ggplot2) 
library(ggthemes)
library(table1)
library(DataExplorer)
```
Load Theme for plots

```{r}
theme_set(theme_fivethirtyeight())
theme_update(axis.title = element_text()) #the default for fivethirtyeight is to not show axis labels, this removes that default so we can choose to specify and display axis titles
theme_update(plot.title = element_text(hjust = 0.5)) # changing default to center all titles
```

Load Data from Kaggle

```{r load data}
titanic = read.csv("https://raw.githubusercontent.com/rickfontenot/Titanic/main/train.csv", header = TRUE)
competition = read.csv("https://raw.githubusercontent.com/rickfontenot/Titanic/main/test.csv", header = TRUE)
gender.sample = read.csv("https://raw.githubusercontent.com/rickfontenot/Titanic/main/gender_submission.csv", header = TRUE)


```

Investigate NA values to determine what needs resolution

```{r}
aggr_plot <- aggr(titanic, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(titanic), cex.axis=.7, gap=3, ylab=c("Percent data missing","Combinations Missing"))

marginplot(titanic[c(4,5)])
view(titanic)
#20% of observations missing data for Age, no other missing values

# NOTE Cabin is missing for ~80% observations too, it's just blank instead on NA

aggr_plot <- aggr(competition, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(competition), cex.axis=.7, gap=3, ylab=c("Percent data missing","Combinations Missing"))

marginplot(competition[c(4,5)])
view(competition)
#20% of observations missing data for Age, 2.4% missing Fare, no other missing values
```
Explore Categorical Variables vs Survival

```{r}
str(titanic)
titanic$Sex <- as.factor(titanic$Sex)
titanic$Embarked <- as.factor(titanic$Embarked)
titanic$Age <- as.integer(titanic$Age)
titanic$Fare <- as.integer(titanic$Fare)

#Class 1 significant higher survival, Class 3 significant lower survival, Class 2 no significant difference
Pclass = table(titanic$Survived, titanic$Pclass)
mosaicplot(Pclass, shade = TRUE, las=2, main = "Pclass", pop = FALSE)
Pclasschisq <- chisq.test(Pclass) 
Pclasschisq #p-value = 2.2e-16

#Significant higher survival for Female
Sex = table(titanic$Survived, titanic$Sex)
mosaicplot(Sex, shade = TRUE, las=2, main = "Sex", pop = FALSE)
Sexchisq <- chisq.test(Sex) 
Sexchisq #p-value = 2.2e-16

#Note SibSp=1 is significant, other values are not, is there something to 1? Create new feature?
SibSp = table(titanic$Survived, titanic$SibSp)
mosaicplot(SibSp, shade = TRUE, las=2, main = "SibSp", pop = FALSE)
SibSpchisq <- chisq.test(SibSp) 
SibSpchisq #p-value = 1.559e-6

#Note Parch=1 is significant, other values are not, is there something to 1? Create new feature for Alone?
Parch = table(titanic$Survived, titanic$Parch)
mosaicplot(Parch, shade = TRUE, las=2, main = "Parch", pop = FALSE)
Parchchisq <- chisq.test(Parch) 
Parchchisq #p-value = 9.7e-5

#Note Too many values for Cabin, need to study distribution, create new feature
Cabin = table(titanic$Survived, titanic$Cabin)
mosaicplot(Cabin, shade = TRUE, las=2, main = "Cabin", pop = FALSE)
Cabinchisq <- chisq.test(Cabin) 
Cabinchisq #p-value = 1.6e-6

#Note Significant higher survival for Embarking at C, S may have small difference but not as significant
Embarked = table(titanic$Survived, titanic$Embarked)
mosaicplot(Embarked, shade = TRUE, las=2, main = "Embarked", pop = FALSE)
Embarkedchisq <- chisq.test(Embarked) 
Embarkedchisq #p-value = 1.6e-6

```
Explore Numerical Variables vs. Survival

```{r}
str(titanic)

Numerical <- titanic %>% select(PassengerId, Age, SibSp, Parch, Fare, Survived)

Numerical_long <- gather(Numerical, key="variable", value="value",-c(PassengerId,Survived))

stat.test <- Numerical_long %>%
  group_by(variable) %>%
  t_test(value ~ Survived) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()

#Sort and display numerical variables with highest significant difference by Attrition
stat.test <- stat.test %>% select(p, everything())
stat.test <- stat.test[order(stat.test$p),] 
stat.test

#Fare p-value=2.62e-11  -> plot survival vs. Fare
#Parch p-value=1.34e-2 -> plot survival vs. Parch
#Age p-value = 4e-2 -> plot survival vs. Age
#SibSp p-value = 0.23 but from chi square above know that SibSp=1 is significant

```

Explore the "Fare" Variable

```{r fare}

titanic %>% ggplot(aes(x = Fare)) + geom_histogram(binwidth=10) + ggtitle("Fare") + xlab("Fare") + ylab("Count")

Fare <- titanic %>% select(Survived,Fare) %>% group_by(Fare) %>% summarise(Survival=mean(Survived),count=n())
plot(Fare$Fare,Fare$Survival)

titanic["FareBin"] <- ifelse(titanic$Fare < 6, "00-05",
                             ifelse((titanic$Fare > 5 & titanic$Fare <11), "06-10",
                             ifelse((titanic$Fare > 10 & titanic$Fare <16), "11-15",
                             ifelse((titanic$Fare > 15 & titanic$Fare <21), "16-20",
                             ifelse((titanic$Fare > 20 & titanic$Fare <31), "21-30",
                             ifelse((titanic$Fare > 30 & titanic$Fare <41), "31-40",
                             ifelse((titanic$Fare > 40 & titanic$Fare <51), "41-50",
                             ifelse((titanic$Fare > 50 & titanic$Fare <61), "51-60",
                             ifelse((titanic$Fare > 60 & titanic$Fare <71), "61-70",
                             ifelse((titanic$Fare > 70 & titanic$Fare <81), "71-80",
                             ifelse((titanic$Fare > 80 & titanic$Fare <91), "81-90",
                             ifelse((titanic$Fare > 90 & titanic$Fare <100), "91-99",
                                                                        "99+"))))))))))))

Fare2 <- titanic %>% select(Survived,FareBin) %>% group_by(FareBin) %>% summarise(Survival=mean(Survived),count=n())
Fare2 %>% ggplot(aes(x = FareBin, y=Survival)) + geom_col() + ggtitle("Fare") + xlab("Fare") + ylab("Count")

plot(titanic$Fare,titanic$Pclass)
plot(titanic$Fare,titanic$Age)
plot(titanic$Fare,titanic$Sex)
plot(titanic$Fare,titanic$Embarked)

titanic %>% select(Fare, Embarked)  %>% ggplot(aes(x=Embarked,y=Fare)) + 
  geom_boxplot()+ labs(title= "Fare by Embark Location" , x = "Embarked", y= "Fare")

titanic %>% select(Fare, Pclass)  %>% ggplot(aes(x=Pclass,y=Fare)) + 
  geom_boxplot()+ coord_flip()+ labs(title= "Fare by Pclass" , x = "Pclass", y= "Fare")

titanic %>% select(Fare, Sex)  %>% ggplot(aes(x=Sex,y=Fare)) + 
  geom_boxplot()+ labs(title= "Fare by Gender" , x = "Gender", y= "Fare")

FareBin = table(titanic$Survived, titanic$FareBin)
mosaicplot(FareBin, shade = TRUE, las=2, main = "FareBin", pop = FALSE)
FareBinchisq <- chisq.test(FareBin) 
FareBinchisq #p-value = 2.2e-16 this is lower for binned Fare than for continuous raw variable

#Fare is for whole booking group, adjust to individual guest fare price
titanic["AdjFare"] = titanic$Fare/(titanic$Parch + titanic$SibSp + 1)

AdjFare = table(titanic$Survived, titanic$AdjFare)
mosaicplot(AdjFare, shade = TRUE, las=2, main = "AdjFare", pop = FALSE)
AdjFarechisq <- chisq.test(AdjFare) 
AdjFarechisq #p-value = 4.7e-12


titanic %>% ggplot(aes(x = AdjFare)) + geom_histogram(binwidth=5) + ggtitle("Fare per person") + xlab("Fare") + ylab("Count")

AdjFare <- titanic %>% select(Survived,AdjFare) %>% group_by(AdjFare) %>% summarise(Survival=mean(Survived),count=n())
plot(AdjFare$AdjFare,AdjFare$Survival)

titanic["AdjFareBin"] <- ifelse(titanic$AdjFare < 6, "00-05",
                             ifelse((titanic$AdjFare > 5 & titanic$AdjFare <11), "06-10",
                             ifelse((titanic$AdjFare > 10 & titanic$AdjFare <16), "11-15",
                             ifelse((titanic$AdjFare > 15 & titanic$AdjFare <21), "16-20",
                             ifelse((titanic$AdjFare > 20 & titanic$AdjFare <31), "21-30",
                             ifelse((titanic$AdjFare > 30 & titanic$AdjFare <41), "31-40",
                             ifelse((titanic$AdjFare > 40 & titanic$AdjFare <51), "41-50",
                             ifelse((titanic$AdjFare > 50 & titanic$AdjFare <61), "51-60",
                             ifelse((titanic$AdjFare > 60 & titanic$AdjFare <71), "61-70",
                             ifelse((titanic$AdjFare > 70 & titanic$AdjFare <81), "71-80",
                             ifelse((titanic$AdjFare > 80 & titanic$AdjFare <91), "81-90",
                             ifelse((titanic$AdjFare > 90 & titanic$AdjFare <100), "91-99",
                                                                        "99+"))))))))))))

Fare2 <- titanic %>% select(Survived,AdjFareBin) %>% group_by(AdjFareBin) %>% summarise(Survival=mean(Survived),count=n())
Fare2 %>% ggplot(aes(x = AdjFareBin, y=Survival)) + geom_col() + ggtitle("FareBin per person") + xlab("AdjFare") + ylab("Count")

plot(titanic$AdjFare,titanic$Pclass)
plot(titanic$AdjFare,titanic$Age)
plot(titanic$AdjFare,titanic$Sex)
plot(titanic$AdjFare,titanic$Embarked)

titanic %>% select(AdjFare, Embarked)  %>% ggplot(aes(x=Embarked,y=AdjFare)) + 
  geom_boxplot()+ labs(title= "AdjFare by Embark Location" , x = "Embarked", y= "AdjFare")

titanic %>% select(AdjFare, Pclass)  %>% ggplot(aes(x=Pclass,y=AdjFare)) + 
  geom_boxplot()+ coord_flip()+ labs(title= "AdjFare by Pclass" , x = "Pclass", y= "AdjFare")

titanic %>% select(AdjFare, Sex)  %>% ggplot(aes(x=Sex,y=AdjFare)) + 
  geom_boxplot()+ labs(title= "AdjFare by Gender" , x = "Gender", y= "AdjFare")

AdjFareBin = table(titanic$Survived, titanic$AdjFareBin)
mosaicplot(AdjFareBin, shade = TRUE, las=2, main = "AdjFareBin", pop = FALSE)
AdjFareBinchisq <- chisq.test(AdjFareBin) 
AdjFareBinchisq #p-value = 1.27e-14 this is higher than the non-adjusted Fare Bin
```

Explore Name and parse into titles, separate by age, do matching surnames equal families

```{r name}
titanic$Title <- substring(titanic$Name,regexpr(",",titanic$Name)+2,regexpr("\\.",titanic$Name)-1)
titanic$Surname <- substring(titanic$Name,0,regexpr(",",titanic$Name)-1)

Title = table(titanic$Survived, titanic$Title)
mosaicplot(Title, shade = TRUE, las=2, main = "Title", pop = FALSE)
Titlechisq <- chisq.test(Title) 
Titlechisq #p-value = 2.2e-16

titanic$Gender[titanic$Title %in% c("Capt","Don","Major","Col","Rev","Dr","Sir","Mr","Jonkheer")] <- "man"
titanic$Gender[titanic$Title %in% c("Dona","the Countess","Mme","Mlle","Ms","Miss","Lady","Mrs")] <- "woman"
titanic$Gender[titanic$Title %in% c("Master")] <- "boy"
view(titanic)

titanic %>% ggplot(aes(x = Age)) + geom_histogram(binwidth=2) + ggtitle("Age") + xlab("Age") + ylab("Count") + facet_wrap(~Gender,ncol=1)

titanic$Gender[titanic$Gender=="woman" & titanic$Age<11] <- "girl"

titanic %>% ggplot(aes(x = Age)) + geom_histogram(binwidth=2) + ggtitle("Age") + xlab("Age") + ylab("Count") + facet_wrap(~Gender,ncol=1)

Gender = table(titanic$Survived, titanic$Gender)
mosaicplot(Gender, shade = TRUE, las=2, main = "Title", pop = FALSE)
Genderchisq <- chisq.test(Gender) 
Genderchisq #p-value = 2.2e-16

titanic %>% ggplot(mapping=aes(y=Sex, x=Gender, color=Survived)) + geom_point(size=1.5, position="jitter")+  labs(title="Sex vs Gender")
```

Explore Ticket and parse into Prefix vs number, look for duplicates, do duplicates equal family members?

```{r ticket}
titanic$Ticket_Pre <- sapply(titanic$Ticket, FUN=function(x) {ifelse(str_detect(x, " "),str_split(x, " ")[[1]][1],'None')})
titanic$Ticket_Pre <- as.factor(str_to_upper(str_replace_all(titanic$Ticket_Pre, "[./]", "")))
titanic$Ticket_Num <- sapply(titanic$Ticket, FUN=function(x) {ifelse(str_detect(x, " "),str_split(x, " ")[[1]][2], as.character(x))})

Ticket = table(titanic$Survived, titanic$Ticket)
mosaicplot(Ticket, shade = TRUE, las=2, main = "Ticket", pop = FALSE)
Ticketchisq <- chisq.test(Ticket) 
Ticketchisq #p-value = 0.01 too many values not significant

Ticket_Pre = table(titanic$Survived, titanic$Ticket_Pre)
mosaicplot(Ticket_Pre, shade = TRUE, las=2, main = "Ticket_Pre", pop = FALSE)
Ticket_Prechisq <- chisq.test(Ticket) 
Ticket_Prechisq #p-value = 0.01 but prefix PC does show significance

titanic["PCticket"] <- "NO"
titanic$PCticket[titanic$Ticket_Pre=="PC"] <- "YES"

PCticket = table(titanic$Survived, titanic$PCticket)
mosaicplot(PCticket, shade = TRUE, las=2, main = "PCticket", pop = FALSE)
PCticketchisq <- chisq.test(PCticket) 
PCticketchisq #p-value = 2.11e-5
```

Explore Cabin and parse the deck vs. the room number

```{r cabin}
titanic$Deck <- substr(titanic$Cabin, 1, 1)
titanic$Deck[titanic$Deck == ""] <- "Unkown"

Cabin = table(titanic$Survived, titanic$Cabin)
mosaicplot(Cabin, shade = TRUE, las=2, main = "Cabin", pop = FALSE)
Cabinchisq <- chisq.test(Cabin) 
Cabinchisq #p-value = 1.6e-06 significant but many unique values based on room numbers

Deck = table(titanic$Survived, titanic$Deck)
mosaicplot(Deck, shade = TRUE, las=2, main = "Deck", pop = FALSE)
Deckchisq <- chisq.test(Deck) 
Deckchisq #p-value = 2.2e-16 
```

Explore Age Variable

```{r age}
titanic %>% ggplot(aes(x = Age)) + geom_histogram(binwidth=5) + ggtitle("Age") + xlab("Age") + ylab("Count")

Age <- titanic %>% select(Survived,Age) %>% group_by(Age) %>% summarise(Survival=mean(Survived),count=n())
plot(Age$Age,Age$Survival)

titanic["AgeBin"] <- ifelse(titanic$Age < 6, "00-05",
                             ifelse((titanic$Age > 5 & titanic$Age <11), "06-10",
                             ifelse((titanic$Age > 10 & titanic$Age <16), "11-15",
                             ifelse((titanic$Age > 15 & titanic$Age <21), "16-20",
                             ifelse((titanic$Age > 20 & titanic$Age <31), "21-30",
                             ifelse((titanic$Age > 30 & titanic$Age <41), "31-40",
                             ifelse((titanic$Age > 40 & titanic$Age <51), "41-50",
                             ifelse((titanic$Age > 50 & titanic$Age <61), "51-60",
                             ifelse((titanic$Age > 60 & titanic$Age <71), "61-70",
                             ifelse((titanic$Age > 70 & titanic$Age <81), "71-80",
                             ifelse((titanic$Age > 80 & titanic$Age <91), "81-90",
                             ifelse((titanic$Age > 90 & titanic$Age <100), "91-99",
                                                                        "99+"))))))))))))

Age2 <- titanic %>% select(Survived,AgeBin) %>% group_by(AgeBin) %>% summarise(Survival=mean(Survived),count=n())
Age2 %>% ggplot(aes(x = AgeBin, y=Survival)) + geom_col() + ggtitle("Fare") + xlab("Fare") + ylab("Survival Rate")

AgeBin = table(titanic$Survived, titanic$AgeBin)
mosaicplot(AgeBin, shade = TRUE, las=2, main = "AgeBin", pop = FALSE)
AgeBinchisq <- chisq.test(FareBin) 
AgeBinchisq #p-value = 2.2e-16 much lower thn continuous variable, Bin under 5years is significant
```
Explore Regression for Age to replace missing values

```{r age regression}

#Split Train & Test
Age_cor <- titanic %>% select(-PassengerId,-Name,-AgeBin,-Ticket,-Cabin)
trainIndices = sample(seq(1:length(Age_cor$Age)),round(.7*length(Age_cor$Age)))
trainAge_cor = Age_cor[trainIndices,]
testAge_cor = Age_cor[-trainIndices,]

#Build Model with all variables to create baseline, then use base model for selection methods
model <- lm(Age ~ ., data = trainAge_cor)
summary(model)

#Variable Selection Methods

#Forward step selection 6 parameters included, RMSE = 12, Adj. R2=0.34
step.forward <- ols_step_forward_p(model)
step.forward

#Backward step selection 3 parameters removed, RMSE = 12, Adj. R2=0.34
step.backward <- ols_step_backward_p(model)
step.backward

#Decision Tree to predict Age
library(rpart)
#install.packages("rattle")
library(rattle)
library(rpart.plot)

tree.age = rpart(Age~., data=trainAge_cor)
print(tree.age$cptable)

fancyRpartPlot(tree.age)

cp <- min(tree.age$cptable[,1])
pruned.tree.age <- prune(tree.age, cp=cp)

testAge_cor_NAfilter <- testAge_cor %>% filter(Age != "NA")
tree.age.predict <- predict(pruned.tree.age, testAge_cor_NAfilter, type="vector")
str(tree.age.predict)
str(testAge_cor_NAfilter$Age)
confusionMatrix(tree.age.predict, testAge_cor_NAfilter$Age)

tree.age.predict.int <- as.integer(tree.age.predict)
str(tree.age.predict.int)
str(testAge_cor_NAfilter$Age)
confusionMatrix(tree.age.predict.int, testAge_cor_NAfilter$Age)

plot(tree.age.predict.int, testAge_cor_NAfilter$Age)
```



Explore Decision Tree for AgeBin to replace missing values

```{r age regression}

#Split Train & Test
Age_cor <- titanic %>% select(-PassengerId,-Name,-Age,-Ticket,-Cabin) %>% filter(AgeBin != "NA")
Age_cor$AgeBin <- as.factor(Age_cor$AgeBin)
trainIndices = sample(seq(1:length(Age_cor$Age)),round(.7*length(Age_cor$Age)))
trainAge_cor = Age_cor[trainIndices,]
testAge_cor = Age_cor[-trainIndices,]

#Decision Tree to predict Age
library(rpart)
#install.packages("rattle")
library(rattle)
library(rpart.plot)

tree.age = rpart(AgeBin~., data=trainAge_cor)
print(tree.age$cptable)

fancyRpartPlot(tree.age)

cp <- min(tree.age$cptable[,1])
pruned.tree.age <- prune(tree.age, cp=cp)

tree.age.predict <- predict(pruned.tree.age, testAge_cor, type="class")
str(tree.age.predict)
str(testAge_cor$AgeBin)
confusionMatrix(tree.age.predict, testAge_cor$AgeBin)

plot(tree.age.predict.int, testAge_cor$AgeBin)
```


Explore Correlations and relationships between variables

```{r corr-EDA}
#Create function to summarize the most significant correlations since there are too many variables for one plot
corr_simple <- function(data=df,sig=0.5){
  #convert data to numeric in order to run correlations
  #convert to factor first to keep the integrity of the data - each value will become a number rather than turn into NA
  df_cor <- data %>% mutate_if(is.character, as.factor)
  df_cor <- df_cor %>% mutate_if(is.factor, as.numeric)
  #run a correlation and drop the insignificant ones
  corr <- cor(df_cor)
  #prepare to drop duplicates and correlations of 1     
  corr[lower.tri(corr,diag=TRUE)] <- NA 
  #drop perfect correlations
  corr[corr == 1] <- NA 
  #turn into a 3-column table
  corr <- as.data.frame(as.table(corr))
  #remove the NA values from above 
  corr <- na.omit(corr) 
  #select significant values  
  corr <- subset(corr, abs(Freq) > sig) 
  #sort by highest correlation
  corr <- corr[order(-abs(corr$Freq)),] 
  #print table
  print(corr)
  #turn corr back into matrix in order to plot with corrplot
  mtx_corr <- reshape2::acast(corr, Var1~Var2, value.var="Freq")
  
  #plot correlations visually
  corrplot(mtx_corr, is.corr=FALSE, tl.col="black", na.label=" ")
}
corr_simple(titanic)

#High Correlations between:
  
```

Age vs Class relationship to Survival

```{r age*class}
titanic["AgeClass"]=titanic$Age * titanic$Pclass

stat.test <- titanic %>%
  t_test(AgeClass ~ Survived) %>%
  adjust_pvalue(method = "BH") %>%
  add_significance()

#Sort and display numerical variables with highest significant difference by Attrition
stat.test <- stat.test %>% select(p, everything())
stat.test <- stat.test[order(stat.test$p),] 
stat.test #p-value 8.67e-22 extremely low, the interaction between Age & Class high impact on survival

titanic %>% ggplot(mapping=aes(y=Age, x=Pclass, color=Survived)) + geom_point(size=1.5, position="jitter")+  labs(title="Class vs Age")
```
Explore feature for people traveling alone using Parch and SibSp

```{r alone}

titanic %>% ggplot(mapping=aes(y=Parch, x=SibSp, color=Survived)) + geom_point(size=1.5, position="jitter")+  labs(title="Parch vs SibSp")

ParchSibsp = table(titanic$Parch, titanic$SibSp)

titanic["TravelGroup"] = titanic$Parch + titanic$SibSp

TravelGroup = table(titanic$Survived, titanic$TravelGroup)
mosaicplot(TravelGroup, shade = TRUE, las=2, main = "TravelGroup", pop = FALSE)
TravelGroupchisq <- chisq.test(TravelGroup) 
TravelGroupchisq #p-value = 3.58e-14 lower survival alone, higher survival for 1,2,3, no difference on others

titanic["Alone"] = "NO"
titanic$Alone[titanic$TravelGroup==0] <- "YES"

Alone = table(titanic$Survived, titanic$Alone)
mosaicplot(Alone, shade = TRUE, las=2, main = "Alone", pop = FALSE)
Alonechisq <- chisq.test(Alone) 
Alonechisq #p-value = 1.973e-09

titanic["GroupSize"] = "Alone"
titanic$GroupSize[titanic$TravelGroup>0 & titanic$TravelGroup<4] <- "Small"
titanic$GroupSize[titanic$TravelGroup>3] <- "Large"

GroupSize = table(titanic$Survived, titanic$GroupSize)
mosaicplot(GroupSize, shade = TRUE, las=2, main = "GroupSize", pop = FALSE)
GroupSizechisq <- chisq.test(GroupSize) 
GroupSizechisq #p-value = 2.2e-16

```

Decision Tree model for Survival

```{r decision tree}
#Decision Tree to predict Survival
library(rpart)
#install.packages("rattle")
library(rattle)
library(rpart.plot)

features1 <- c("Survived","AgeClass","AgeBin","Pclass","Sex","FareBin","Deck","GroupSize","Embarked","PCticket")

#Split Train & Test
TreeInput <- titanic %>% select(Survived,AgeClass,AgeBin,Pclass,Sex,FareBin,Deck,GroupSize,Embarked,PCticket) %>% filter(AgeBin != "NA")
TreeInput$AgeBin <- as.factor(TreeInput$AgeBin)

view(TreeInput)

trainIndices = sample(seq(1:length(TreeInput$Survived)),round(.7*length(TreeInput$Survived)))
trainTreeInput = TreeInput[trainIndices,]
testTreeInput = TreeInput[-trainIndices,]


tree.survived = rpart(Survived~., data=trainTreeInput)
print(tree.survived$cptable)

fancyRpartPlot(tree.survived)

cp <- min(tree.survived$cptable[,1])
pruned.tree.survived <- prune(tree.survived, cp=cp)

fancyRpartPlot(pruned.tree.survived)

testTreeInput["SurvivalProbability"] <- predict(pruned.tree.survived, testTreeInput, type="vector")

testTreeInput["Prediction"] = 0
testTreeInput$Prediction[testTreeInput$SurvivalProbability>0.5] = 1
testTreeInput$Prediction=as.factor(testTreeInput$Prediction)
testTreeInput$Survived=as.factor(testTreeInput$Survived)

confusionMatrix(testTreeInput$Prediction, testTreeInput$Survived)
#Accuracy 80.84%

plot(testTreeInput$Prediction, testTreeInput$Survived)

testTreeInput %>% ggplot(aes(x=as.character(Survived),y=Prediction)) + 
  geom_boxplot()+ labs(title= "Prediction vs Survived" , x = "Survived", y= "Prediction")

testTreeInput %>% ggplot(aes(x=as.character(Survived),y=Prediction)) + 
  geom_boxplot()+ labs(title= "Prediction vs Survived" , x = "Survived", y= "Prediction")

testTreeInput %>% ggplot(aes(x = SurvivalProbability, fill=Survived)) + geom_histogram() +
ggtitle("Probability Survived by Actual Survival") + xlab("Probability") + ylab("Count") +facet_wrap(~Survived,ncol=1)
```

knn model

```{r knn}
features1 <- c("Survived","AgeClass","AgeBin","Pclass","Sex","FareBin","Deck","GroupSize","Embarked","PCticket")

#Split Train & Test
knn.input <- titanic %>% select(features1) %>% filter(AgeBin != "NA")
#Convert Categorical to numeric
knn.input$AgeBin <- as.factor(knn.input$AgeBin)
knn.input$AgeBin <- as.numeric(knn.input$AgeBin)
knn.input$Sex <- as.factor(knn.input$Sex)
knn.input$Sex <- as.numeric(knn.input$Sex)
knn.input$FareBin <- as.factor(knn.input$FareBin)
knn.input$FareBin <- as.numeric(knn.input$FareBin)
knn.input$Deck <- as.factor(knn.input$Deck)
knn.input$Deck <- as.numeric(knn.input$Deck)
knn.input$GroupSize <- as.factor(knn.input$GroupSize)
knn.input$GroupSize <- as.numeric(knn.input$GroupSize)
knn.input$Embarked <- as.factor(knn.input$Embarked)
knn.input$Embarked <- as.numeric(knn.input$Embarked)
knn.input$PCticket <- as.factor(knn.input$PCticket)
knn.input$PCticket <- as.numeric(knn.input$PCticket)


trainIndices = sample(seq(1:length(knn.input$Survived)),round(.7*length(knn.input$Survived)))
knn.train = knn.input[trainIndices,]
knn.test = knn.input[-trainIndices,]

knnpredictions = knn(knn.train[,2:10],knn.test[,2:10],knn.train$Survived,k=5,prob = TRUE)
table(knn.test$Survived,knnpredictions)
confusionMatrix(table(knn.test$Survived,knnpredictions))
    
```

Naive Bayes model

```{r}

features2 = c("AgeClass","AgeBin","Pclass","Sex","FareBin","Deck","GroupSize","Embarked","PCticket")


model = naiveBayes(knn.train[,features2],as.factor(knn.train$Survived),laplace=1)
confusionMatrix(table(predict(model,knn.test[,features2]),knn.test$Survived))

```

Logistic Regression Model 

```{r logistic regression}
logit.train <- knn.train
logit.test <- knn.test

model <- glm(Survived ~.,family=binomial(link='logit'),data=logit.train)

summary(model)

logit.train$SurvivalProbability <- predict(model, newdata = logit.train, type = "response")

logit.train["Prediction"] = 0
logit.train$Prediction[logit.train$SurvivalProbability>0.5] = 1
logit.train$Prediction=as.factor(logit.train$Prediction)
logit.train$Survived=as.factor(logit.train$Survived)

confusionMatrix(logit.train$Prediction, logit.train$Survived)
#80% accuracy on training set used in model

logit.test$SurvivalProbability <- predict(model, newdata = logit.test, type = "response")

logit.test["Prediction"] = 0
logit.test$Prediction[logit.test$SurvivalProbability>0.5] = 1
logit.test$Prediction=as.factor(logit.test$Prediction)
logit.test$Survived=as.factor(logit.test$Survived)

confusionMatrix(logit.test$Prediction, logit.test$Survived)
#79.91% accuracy on training set used in model

```

Analyze misclassifications for logistic regression model and look for variables to optimize

```{r optimize}

#Create category for prediction accuracy
logit.test <- logit.test %>% 
  mutate(category = case_when(
    Survived == 0 & Prediction == 0 ~ "True Negative",
    Survived == 1 & Prediction == 1 ~ "True Positive",
    Survived == 1 & Prediction == 0 ~ "False Negative",
    Survived == 0 & Prediction == 1 ~ "False Positive",
    TRUE ~ "other"
  )
)

logit.test %>% group_by(category) %>% summarise(n = n())
#Summary matches confusion matrix:
#False Negative	21			
#False Positive	22			
#True Negative	103			
#True Positive	68	

#install.packages("esquisse")
#remotes::update_packages("htmltools")
#library(esquisse)

#features2 = c("AgeClass","AgeBin","Pclass","Sex","FareBin","Deck","GroupSize","Embarked","PCticket")

AgeBincategory = table(logit.test$AgeBin, logit.test$category)
mosaicplot(AgeBincategory, shade = TRUE, las=2, main = "AgeBincategory", pop = FALSE)
AgeBincategorychisq <- chisq.test(AgeBincategory) 
AgeBincategorychisq #Significant higher false classifications for AgeBins 2,3,6

Pclasscategory = table(logit.test$Pclass, logit.test$category)
mosaicplot(Pclasscategory, shade = TRUE, las=2, main = "Pclasscategory", pop = FALSE)
Pclasscategorychisq <- chisq.test(Pclasscategory) 
Pclasscategorychisq #not significant for flase classifications

Sexcategory = table(logit.test$Sex, logit.test$category)
mosaicplot(Sexcategory, shade = TRUE, las=2, main = "Sexcategory", pop = FALSE)
Sexcategorychisq <- chisq.test(Sexcategory) 
Sexcategorychisq #Higher false positives for sex=1

FareBincategory = table(logit.test$FareBin, logit.test$category)
mosaicplot(FareBincategory, shade = TRUE, las=2, main = "FareBincategory", pop = FALSE)
FareBincategorychisq <- chisq.test(FareBincategory) 
FareBincategorychisq #Higher false for FareBin=1,4,6 especially look at 1

logit.test %>% group_by(category,AgeBin) %>% summarise(n = n())
view(logit.test)
view(knn.test)
```